---
title: "Similarity Judgment Task Verbs2 (Phys-Psych): Analysis"
author: "morgan moyer"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE, warning=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(lme4)
library(lmerTest)
library(multcomp) 
library(stringr)
library(textstem)
library(tidyverse)
library(broom.mixed)  # for tidy()
library(ggeffects)
library(lattice)

theme_set(theme_bw())
cbPalette <- c("#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00")
this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(this.dir)
source("../../helpers.R")

d <- read.csv("../data/processed.csv")
d["WordPair"] <- paste(d$Word1,d$Word2,sep="-")
sim <- read.csv("../data/pilot_similarity.csv")
sim["WordPair"] <- paste(sim$Word1,sim$Word2,sep="-")
sim <- sim[,c("WordPair","W2VSimilarity","CosineSimilarity")]

d <- merge(d,sim, by="WordPair")

# d <- d %>% 
#   filter(!FeatureMatch == "SelfPair")
```


```{r}
names(d)
dcen <- d %>% 
  mutate(WordPair = as.factor(WordPair),
         ID.Ibex = as.factor(ID.Ibex),
         FeatureMatch = as.factor(FeatureMatch),
         cResponse = as.numeric(Response)-mean(as.numeric(Response)))

m <- lmer(cResponse ~ FeatureMatch + (1|WordPair) + (1+FeatureMatch|ID.Ibex), data = dcen)
summary(m)
```

### Plot fixed effect estimates with confidence intervals

```{r}
# Tidy the model output
coefs <- tidy(m, effects = "fixed")
# Reorder by effect size for prettier plotting
coefs$term <- factor(coefs$term, levels = coefs$term[order(coefs$estimate)])

# Then replot
ggplot(coefs, aes(x = estimate, y = term, color = term)) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = estimate - 1.96 * std.error,
                     xmax = estimate + 1.96 * std.error), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  # theme_minimal(base_size = 14) +
  labs(title = "Fixed Effect Estimates from Mixed Model",
       x = "Estimate (95% CI)", y = NULL) +
  theme(legend.position="none")

```


### Plot predicted values vs. predictors
You can create a dataset of predicted values from your model and plot how the response changes with predictors:

1. Create a new dataset with combinations of predictors
- Generates all combinations of levels of FeatureMatch (a factor) and a sequence of W2VSimilarity values from -0.01 to 1 (100 points).
- Because: You want to predict your response across a continuous range of W2VSimilarity for each FeatureMatch group.
```{r}
newdata <- expand.grid(
  FeatureMatch = levels(dcen$FeatureMatch))
```

2. Get predicted values and standard errors from the model
```{r}
pred <- predict(m, newdata = newdata, re.form = NA, se.fit = TRUE)
```

3. Add predictions and confidence intervals to newdata
```{r}
newdata$fit <- pred$fit
newdata$se <- pred$se.fit
newdata <- newdata %>% 
  mutate(
    conf.low = fit - 1.96 * se,
    conf.high = fit + 1.96 * se
  )
```

4. Plot predicted lines with ribbons for confidence intervals
```{r}
ggplot(newdata, aes(x = reorder(FeatureMatch,fit), y = fit, color = FeatureMatch, fill = FeatureMatch)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  # theme_minimal() +
  labs(x = "FeatureMatch", y = "Predicted Response", color = "FeatureMatch", fill = "FeatureMatch") +
  theme(legend.position="none")

```


## You can visualize individual random effect estimates:


What you get from the plot:
A visual summary of the variability in random effects across levels of your grouping factors.

Helps you identify:

Which groups (subjects/items) deviate strongly from the average effect,

How precise these deviations are,

Whether there are outliers or patterns in random effects.

Example interpretation:
If a subject’s random intercept dot is far from zero with narrow error bars, that subject consistently differs from the average intercept with high certainty.

```{r}
dotplot(ranef(m, condVar=TRUE))

```

## Check model diagnostics
For residuals, fitted vs residuals, etc.:

```{r}
plot(m)
```



# Does Similarity predict or interact with FeatureMatch?

## Cosine Similarity from Transformer Models
- No.
```{r}
m.full <- lmer(cResponse ~ FeatureMatch*CosineSimilarity + (1|WordPair) + (1|ID.Ibex), data = dcen)
summary(m.full)
```

## Does adding Cosine Similarity explain any more variance in the data?
- no
```{r}
m.s <- lmer(cResponse ~ FeatureMatch + (1|WordPair) + (1|ID.Ibex), data = dcen)

anova(m.s, m.full, REML=FALSE)
```

### Plot fixed effect estimates with confidence intervals

```{r}
# Tidy the model output
coefs <- tidy(m.full, effects = "fixed")

# Reorder by effect size for prettier plotting
coefs$term <- factor(coefs$term, levels = coefs$term[order(coefs$estimate)])

# Then replot
ggplot(coefs, aes(x = estimate, y = term, color = term)) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = estimate - 1.96 * std.error,
                     xmax = estimate + 1.96 * std.error), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  # theme_minimal(base_size = 14) +
  labs(title = "Fixed Effect Estimates from Mixed Model",
       x = "Estimate (95% CI)", y = NULL) +
  theme(legend.position="none")
```



## Word-2-Vec Similarity
- yes some interactions...
1. FeatureMatchMaxMatch:W2VSimilarity             140.280     43.000  41.000   3.262  0.00223 **
2. FeatureMatchSelfPair                            99.667     29.758  42.059   3.349  0.00172 **


```{r}
m.full <- lmer(cResponse ~ FeatureMatch*W2VSimilarity + (1|WordPair) + (1|ID.Ibex), data = dcen)
summary(m)
```

## Does adding Word-2-Vec Similarity explain any more variance in the data?
- Yes
```{r}
anova(m.s, m.full, REML=FALSE)
```

### Plot fixed effect estimates with confidence intervals
```{r}
# Tidy the model output
coefs <- tidy(m.full, effects = "fixed")

# Reorder by effect size for prettier plotting
coefs$term <- factor(coefs$term, levels = coefs$term[order(coefs$estimate)])

# Then replot
ggplot(coefs, aes(x = estimate, y = term, color = term)) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = estimate - 1.96 * std.error,
                     xmax = estimate + 1.96 * std.error), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  # theme_minimal(base_size = 14) +
  labs(title = "Fixed Effect Estimates from Mixed Model",
       x = "Estimate (95% CI)", y = NULL) +
  theme(legend.position="none")

```

### Plot predicted values vs. predictors
You can create a dataset of predicted values from your model and plot how the response changes with predictors:

1. Create a new dataset with combinations of predictors
- Generates all combinations of levels of FeatureMatch (a factor) and a sequence of W2VSimilarity values from -0.01 to 1 (100 points).
- Because: You want to predict your response across a continuous range of W2VSimilarity for each FeatureMatch group.
```{r}
newdata <- expand.grid(
  FeatureMatch = levels(dcen$FeatureMatch),
  W2VSimilarity = seq(-0.01, 1, length.out = 100)
)
```

2. Get predicted values and standard errors from the model
```{r}
pred <- predict(m.full, newdata = newdata, re.form = NA, se.fit = TRUE)
```

3. Add predictions and confidence intervals to newdata
```{r}
newdata$fit <- pred$fit
newdata$se <- pred$se.fit
newdata <- newdata %>% 
  mutate(
    conf.low = fit - 1.96 * se,
    conf.high = fit + 1.96 * se
  )
```

4. Plot predicted lines with ribbons for confidence intervals
```{r}
ggplot(newdata, aes(x = W2VSimilarity, y = fit, color = FeatureMatch, fill = FeatureMatch)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, color = NA) +
  theme_minimal() +
  labs(x = "Word-2-Vec Similarity", y = "Predicted Response", color = "FeatureMatch", fill = "FeatureMatch")
```


### Visualize individual random effect estimates:

```{r}
dotplot(ranef(m.full, condVar=TRUE))

```

### Check model diagnostics
For residuals, fitted vs residuals, etc.:

```{r}
plot(m.full)
```


# Similarity Alone
```{r}
m <- lmer(cResponse ~ W2VSimilarity + (1|WordPair) + (1|ID.Ibex), data = dcen)
summary(m)
```

```{r}
m <- lmer(cResponse ~ CosineSimilarity + (1|WordPair) + (1|ID.Ibex), data = dcen)
summary(m)
```

